---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


[![Travis-CI Build Status](http://badges.herokuapp.com/travis/Trackage/trip?branch=master&env=BUILD_NAME=trusty_release&label=linux)](https://travis-ci.org/Trackage/trip)
[![OSX Build Status](http://badges.herokuapp.com/travis/Trackage/trip?branch=master&env=BUILD_NAME=osx_release&label=osx)](https://travis-ci.org/Trackage/trip)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/mdsumner/trip?branch=master&svg=true)](https://ci.appveyor.com/project/mdsumner/trip)[![Coverage_Status](https://img.shields.io/codecov/c/github/Trackage/trip/master.svg)](https://codecov.io/github/Trackage/trip?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/trip)](https://cran.r-project.org/package=trip)
[![CRAN_Download_Badge](http://cranlogs.r-pkg.org/badges/trip)](https://cran.r-project.org/package=trip)


# Tools for animal track data

The trip package provides functions for accessing and manipulating
spatial data for animal tracking.  Filter for speed and create time
spent plots from animal track data.


## Installing

The package is easily installed from CRAN in R. 

```R
install.packages("trip")
```

## Install dev version

To install the development package from Github: 

```R
remotes::install_github("Trackage/trip")
```
# Development


Currently working on tidying up old daggy ends, see [next release milestones](https://github.com/Trackage/trip/milestone/2). 

Particulary, I want it to be very easy to create trip objects from other forms, either by reading known formats (Argos), coercion from other packages (sp, sf, adehabitatLT, move, etc.), or from raw data with a dplyr/gg aesthetic e.g. this kind of construct from a data frame: 

```R
## NOTE: this is pseudocode, neither works yet
d %>% arrange(id, time) %>% group_by(id) %>%  select(x, y, time, everything()) %>% trip(crs = "+proj=longlat")

## OR this for the same outcome
trip(d, x, y, time, id, crs = "+proj=longlat")
```

We could assume longlat is input if it has sane ranges, much like raster does. 



There is some ongoing integration with the `sf` package, mostly to provide outputs in its
formats. 

We can already convert to POINT: 

```{r coerce-sf-1}
library(trip)
sf::st_as_sf(walrus818)

```
or atomic LINESTRING with time on the MEASURE:

```{r coerce-sf-2}
as(walrus818, "sf")
```
Trip itself would be better implemented on top of the ideas in [the silicate package](https://github.com/hypertidy/silicate/), rather than the twilight zone between MULTIPOINTs and LINESTRINGs as it is now. 


Let me know what you think via [Issues](https://github.com/Trackage/trip/issues). 

## TODO

- **Probability image**.  The SGAT (and tripEstimation) packages have
functions for dealing with spatial track summaries that are
atomized to the level of each time step. There are methods for combining summaries from
multiple tracks and for casting arbitrary durations (by sum) to standard image structures. This would be a good
feature to replace the existing tripGrid function by storing the 
individual grid summaries for each implicit line segment.

- **Coercion to/from other classes** See [spbabel](https://github.com/mdsumner/spbabel). 

- **Validation** Must include a detailed report object of where the problems are, and how to filter/fix/flush them. 
 
 
 
 
Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
