---
title: "The trip package"
author: Michael Sumner
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 7
    fig_height: 7
vignette: >
  %\VignetteIndexEntry{trip}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


# Introduction

Basic use of the trip package.

## Data input and validation

Tracking data is in essence a grouped series of one-dimensional data[^1] records. 

It is *grouped* because we may track more than one object, each with its on
independent time sequence.

It is *one-dimensional* because the topology of the data is a linear stream of
values. It makes sense that time is constantly increasing, and it is the primary
dimension of the process. (Location is often something that we must estimate, but time
is usually directly measured and robust). 

Commonly we have records in a data frame, and the primary workflow is to provide
the `trip()` function with information about the spatial locations, the temporal
data, and the grouping.

The data may contain any other columns, and in general they can be called anything and
be in any order, but the simplest way to create a `trip` is to put the first four columns 
as X, Y, date-time, grouping. 

```{r}
library(trip)
d <- data.frame(x=1:10,y=rnorm(10), tms=Sys.time() + 1:10, id=gl(2, 5))
tr <- trip(d)
summary(tr)
```

(There may be only one group, but we have to be explicit, with a column that identifies at least one group). 


When a print or summary is made the data are presented in terms of their grouping, with some handy summary values. When converting to `sf` or `sp` form as lines these summary values are recorded with each "multi"-line. 

# Simple plotting


To plot a trip we use base graphics in the usual ways. 

```{r}
plot(tr)
lines(tr)
```

The trip object acts as a `sp` data frame of points, but with the underlying grouping as lines when that is relevant. 

```{r}
plot(tr,pch = ".", col = rainbow(nrow(tr)))
lines(tr, col = c("dodgerblue", "firebrick"))
```


# Gridding for time spent

There is a key functionality to determine the *time spent in area* on a grid, by leveraging the `rasterize()` generic function in the raster package. Any raster object may be used, so the specification of pixel area, extent and map projection is up to the user.  (The trip line segments must all fall within the raster). 

```{r}
tg <- rasterize(tr)
plot(tg, col = c("transparent", heat.colors(25)))
```

As with `raster::rasterize()` the `field` used may be chosen, by defaul the time difference between each point in a trip is used, and the final grid contains the sum of these durations. 

There is an older version of this in the `tripGrid.interp` function that uses approrimate methods by allowing interpolation to an 'equal time' step. 

# Reading from Argos files

Service Argos has provide various message formats, and the `readArgos()` function understands some
variants of the 'PRV' form. Multiple files can be provided, and all messages will be normalized and
turned into a trip object with no need for the user to group or clean them in any way. 

```{r read-argos}
argosfile <- 
 system.file("extdata/argos/98feb.dat", package = "trip", mustWork = TRUE)
argos <- readArgos(argosfile) 

summary(argos)
```

Note that the form of the coordinates is native to the PRV file, and in this case contains longitude values that are greater than 180. Some forms of these files provide wrapped forms, but in general the data are
read as-is. 

```{r plot-argos-anti-meridian}
plot(argos, pch = "l", asp = f)
lines(argos)
maps::map("world2", add = TRUE)
sp::degAxis(1)
sp::degAxis(2)
```

# Walrus tracks

Plot the walrus tracks on a map. 

```{r}
data("walrus818")

plot(walrus818, pch = ".")
lines(walrus818)
```


```{r}
library(maptools)
library(rgdal)
data(wrld_simpl)
world <- sp::spTransform(subset(wrld_simpl, coordinates(wrld_simpl)[,2] > 0), proj4string(walrus818))
p <- par(xpd = NA, mar = rep(0.5, 4))
plot(walrus818, pch = ".")
plot(world, add = TRUE, col = "grey")
lines(walrus818)
llgridlines(walrus818); par(p)
```




[^1]: *One-dimensional, are you crazy?* Yes, the measurement process is one-dimensional and that is how we
can arrange the primary data we collect. We collect location (x, y, z), time, and many other variables such as temperature, air pressure, happiness and colour, these are the geometry of our measurements, but the collection itself is very much a one-dimensional topology.
